---
title: "Final Project"
author: "Tsu-yen Su"
output: html_notebook
---

t. 

## Research Question 

How does the performance of individual players on the Denver team vary between regular-season and playoff games?

It's interesting to see how players on the Denver team perform differently in regular-season versus playoff games. It gives us a glimpse into how they handle the pressure and adapt their game, which can help the team make better strategies and decisions during the playoffs.

The origin of the data is the ESPN website, specifically the pages containing the statistics of Denver Nuggets players. The data were collected during the playoff season and regular season of 2022-2023. The purpose of collecting this data is to provide insights into the performance of individual Denver Nuggets players, specifically how they handle pressure and adapt their game. This information can be valuable for the team's coaching staff and management to make better strategies and decisions during the playoffs. The data is collected from ESPN.com.

## Data Processing 

#### Data Intake 

Include R code and written explanation to import your **TWO** data sets.

## (A) .Rmd Line numbers where at least two different data sources are imported:  
```{r}
library(rvest)
```

# Regular Season Player Stats

```{r}
regular_season_url <- "https://www.espn.com/nba/team/stats/_/name/den/season/2023/seasontype/2"
regular_season_table <- regular_season_url %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)
```


# Playoff Season Player Stats


```{r}
playoff_season_url <- "https://www.espn.com/nba/team/stats/_/name/den/season/2023/seasontype/3"
playoff_season_table <- playoff_season_url %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)
```
## (B) .Rmd Line numbers for inspecting data intake:  
#Regular Season
```{r}
library(dplyr)
library(rvest)

regular_season_url <- "https://www.espn.com/nba/team/stats/_/name/den/season/2023/seasontype/2"
regular_season_table <- regular_season_url %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)

glimpse(regular_season_table)

```
#Playoff Season
```{r}
library(dplyr)
library(rvest)

playoff_season_url <- "https://www.espn.com/nba/team/stats/_/name/den/season/2023/seasontype/3"
playoff_season_table <- playoff_season_url %>%
  read_html() %>%
  html_nodes(css = "table") %>%
  html_table(fill = TRUE)

glimpse(playoff_season_table)

```


#### Data Wrangling 

Include R code and written explanation for wangling your data (you can make multiple wrangled data sets). 

##(A) .Rmd Line number(s) for general data wrangling:

# Assuming the desired table is the Second one

```{r}
library(dplyr)
regular_season_df <- regular_season_table[[2]] 
playoff_season_df <- playoff_season_table[[2]] 
```
##(B) .Rmd Line number(s) for a join operation:  

# Combine the data frames
```{r}
regular_season_w_name <- bind_cols(regular_season_df,regular_season_table[[1]])



regular_season_w_name <-
regular_season_w_name %>% 
  relocate(Name)

playoff_season_w_name <- bind_cols(playoff_season_df,playoff_season_table[[1]])



playoff_season_w_name <-
playoff_season_w_name %>% 
  relocate(Name)



combined_df <- full_join(regular_season_w_name, playoff_season_w_name,by="Name")
```
##(C) .Rmd Line number(s) for use of reduction and/or transformation functions

# Extract the top 10 scoring players on both regular season and playoff season
```{r}
library(dplyr)

combined_df <- combined_df %>%
  slice(1:10)




```
## (D) .Rmd Line number(s) for a pivot_wider or pivot_longer operation (or equivalent):
```{r}
library(tidyr)

# Regular Season Player Stats
regular_season_data <- data.frame(
  NAME = c("Nikola Jokic", "Jamal Murray", "Michael Porter Jr.", "Aaron Gordon"),
  GP = c(69, 65, 62, 68),
  PTS = c(24.5, 20.0, 17.4, 16.3),
  REB = c(11.8, 4.0, 5.5, 6.6),
  AST = c(9.8, 6.2, 1.0, 3.0)
)

# Pivot from wide to long format
regular_season_long <- pivot_longer(regular_season_data, cols = c(GP, PTS, REB, AST), names_to = "Statistic", values_to = "Value")

# Output the long format
regular_season_long

```
##(G) .Rmd Line number(s) for use of loops and/or control flow: 
#Total points scored 
```{r}
# Regular Season Player Stats
regular_season_data <- data.frame(
  NAME = c("Nikola Jokic", "Jamal Murray", "Michael Porter Jr.", "Aaron Gordon"),
  GP = c(69, 65, 62, 68),
  PTS = c(24.5, 20.0, 17.4, 16.3),
  REB = c(11.8, 4.0, 5.5, 6.6),
  AST = c(9.8, 6.2, 1.0, 3.0)
)

# Create an empty list to store processed data
processed_data <- list()

# Loop through each player's stats
for (i in 1:nrow(regular_season_data)) {
  player_name <- regular_season_data$NAME[i]
  games_played <- regular_season_data$GP[i]
  points <- regular_season_data$PTS[i]
  rebounds <- regular_season_data$REB[i]
  assists <- regular_season_data$AST[i]
  
  # Perform any desired transformations or calculations
  total_points <- points * games_played
  
  # Store the processed data in a new data frame
  processed_data[[i]] <- data.frame(Player = player_name, TotalPoints = total_points, Rebounds = rebounds, Assists = assists)
}

# Combine all the processed data frames into a single data frame
final_data <- do.call(rbind, processed_data)

# Output the final data frame
final_data

```

## Data Visualization

#### Plot 1

## (A) .Rmd Line number(s) for use of multiple different geoms on a single plot:  
# Plot 1: Average Points per Game During Regular Season


```{r}
# Plot 1: Average Points per Game comparison

library(ggplot2)

plot1 <- ggplot(combined_df, aes(x = reorder(Name, -PTS.x), y = PTS.x)) +
  geom_bar(stat = "identity", fill = "blue") +
  geom_text(aes(label = PTS.x), vjust = -0.5, size = 3.5) +
  labs(title = "Average Points per Game During Regular Season",
       x = "Player", y = "Average Points per Game") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

plot1




```

#### Plot 2:: Average Per Game During Playoff Season

##(B) .Rmd Line number(s) for use of multiple aesthetics: 
```{r}

plot2 <- ggplot(combined_df, aes(x = reorder(Name, -PTS.y), y = PTS.y)) +
  geom_bar(stat = "identity", fill = "green") +
  geom_text(aes(label = PTS.y), vjust = -0.5, size = 3.5) +
  labs(title = "Average Points per Game During Playoff Season",
       x = "Player", y = "Average Points per Game") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```
#### Plot 3:: Total points, rebounds, and assists for each player

##(C) .Rmd Line number(s) for use of layered graphics: 
```{r}
library(ggplot2)

# Plot the total points, rebounds, and assists for each player
ggplot(final_data, aes(x = Player)) +
  geom_col(aes(y = TotalPoints, fill = "Total Points"), width = 0.5, position = "dodge") +
  geom_col(aes(y = Rebounds, fill = "Rebounds"), width = 0.5, position = "dodge") +
  geom_col(aes(y = Assists, fill = "Assists"), width = 0.5, position = "dodge") +
  labs(x = "Player", y = "Value", fill = "Category") +
  scale_fill_manual(values = c("Total Points" = "blue", "Rebounds" = "green", "Assists" = "orange")) +
  theme_minimal()
```


## Conclusion
Based on the analysis of the average points per game for individual players on the Denver team during the regular season and playoff season, several key observations can be made:

Overall Performance: The data shows that the players' average points per game generally increased during the playoff season compared to the regular season. This indicates that the players were able to elevate their performance and contribute more offensively during the high-pressure playoff games.

Jokic Dominance: Nikola Jokic stands out as a highly consistent and impactful player. He averaged 24.5 points per game during the regular season, which increased to an impressive 30 points per game during the playoffs. This suggests that Jokic thrives under pressure and significantly elevates his scoring output in the postseason.

Jamal Murray's Impact: Jamal Murray also demonstrated a significant improvement in his performance during the playoffs. He averaged 20 points per game during the regular season, which notably rose to 26.1 points per game in the postseason. Murray's ability to elevate his scoring showcases his importance as a key contributor during playoff games.

Varied Performances: While Jokic and Murray consistently performed at a high level, there were some players who experienced fluctuations in their performance. Michael Porter Jr. averaged 17.4 points per game during the regular season, which decreased to 13.4 points per game in the playoffs. On the other hand, Aaron Gordon saw a slight increase in his scoring from 16.3 to 13.3 points per game.

In conclusion, the data analysis highlights the varying performances of individual players on the Denver team between the regular season and playoff games. It shows an overall improvement in scoring during the playoffs, with Jokic and Murray emerging as key contributors with significant increases in their average points per game. However, there were some players who experienced fluctuations or a decrease in their scoring output. This analysis provides insights into how players on the Denver team handle the pressure and adapt their game during the playoffs, which can be valuable for the team's coaching staff and management in making strategies and decisions for future playoff campaigns.

Furthermore, the provided data showcases the specific total points, rebounds, and assists for each player:

Player	TotalPoints	Rebounds	Assists
Nikola Jokic	1690.5	11.8	9.8
Jamal Murray	1300.0	4.0	6.2
Michael Porter Jr.	1078.8	5.5	1.0
Aaron Gordon	1108.4	6.6	3.0
These additional statistics provide further context to the players' performances and contribute to the overall understanding of their contributions to the team.

